### 2017-01-05 experiments.tex
added spsa-3 tables

\chapter{Simulation experiments}
\label{sec:expts}
\section{Implementation}
We test the performance of 2RDSA-Unif, 2RDSA-AsymBer, 2SPSA-3 and 2SPSA, with/without improved Hessian estimation. 
2SPSA and 2SPSA-3 algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.

@@

\begin{table}
\centering
 \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-4thf}
\scalebox{0.9}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
&&\\
\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
&&\\
\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
 \bottomrule
\end{tabular}
}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%% Quadratic
\begin{table}
\centering
 \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-quadratic}
\scalebox{0.9}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0062 \pm 0.1164$ & $-0.1229 \pm 0.1374$\\
&&\\
\textbf{2RDSA-Unif} &$0.0485 \pm 0.1465$ & $-0.259 \pm 0.0398$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2564 \pm 0.068$& $\bm{-0.2877 \pm 0.0051}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0785 \pm 0.1178$ & $-0.1716 \pm 0.1339$\\
&&\\
\textbf{2RDSA-Unif} &$0.0326 \pm 0.1599$ & $-0.2672 \pm 0.0299$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2777 \pm 0.0488$& $\bm{-0.2881 \pm 0.0012}$\\
 \bottomrule
\end{tabular}}

@@

%%%%%%%%%%%%%% NMSE for quadratic
\begin{table}
\centering
 \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:nmse-quadratic}
\scalebox{0.9}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
&&\\
\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
&&\\
\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
 \bottomrule
\end{tabular}}
\end{table}

@@

Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively. Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss. The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
 Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
 From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:

>>>>

\chapter{Simulation experiments}
\label{sec:expts}
\section{Implementation}
We test the performance of 2RDSA-Unif, 2RDSA-AsymBer, 2SPSA-3 and 2SPSA, with/without improved Hessian estimation. 
2SPSA and 2SPSA-3 algorithms use Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.

@@

\begin{table}
\centering
 \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-4thf}
\scalebox{0.8}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
&&\\
\textbf{2SPSA-3} & $0.080 \pm 0.0053$ & $0.043 \pm 0.0015$\\
&&\\
\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
&&\\
\textbf{2SPSA-3} & $0.0338 \pm 0.0001$ & $0.0315 \pm 0.0007$\\
&&\\
\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
 \bottomrule
\end{tabular}
}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%% Quadratic
\begin{table}
\centering
 \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-quadratic}
\scalebox{0.8}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0062 \pm 0.1164$ & $-0.1229 \pm 0.1374$\\
&&\\
\textbf{2SPSA-3} & $-0.0161 \pm 0.0955$ & $-0.2748 \pm 0.0629$\\
&&\\
\textbf{2RDSA-Unif} &$0.0485 \pm 0.1465$ & $-0.259 \pm 0.0398$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2564 \pm 0.068$& $\bm{-0.2877 \pm 0.0051}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0785 \pm 0.1178$ & $-0.1716 \pm 0.1339$\\
&&\\
\textbf{2SPSA-3} & $-0.0667 \pm 0.0602$ & $-0.2793 \pm 0.0625$\\
&&\\
\textbf{2RDSA-Unif} &$0.0326 \pm 0.1599$ & $-0.2672 \pm 0.0299$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2777 \pm 0.0488$& $\bm{-0.2881 \pm 0.0012}$\\
 \bottomrule
\end{tabular}}

@@

@@

%%%%%%%%%%%%%% NMSE for quadratic
\begin{table}
\centering
 \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:nmse-quadratic}
%\scalebox{0.8}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
&&\\
\textbf{2SPSA-3} & $0.8378 \pm 0.0179$ & $0.1045 \pm 0.0005$\\
&&\\
\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
&&\\
\textbf{2SPSA-3} & $0.6661 \pm 0.01554$ & $0.0684 \pm 0.0006$\\
&&\\
\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
 \bottomrule
\end{tabular}
\end{table}

Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the four algorithms - 2SPSA, 2SPSA-3, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively. Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss. The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.

 Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$. Figure \ref{fig:QuadnorlossvssimsNoiseless} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$. See Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless}  for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss. 
 From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:

### 2017-01-05 intro.tex
experiments added
added spsa-3 tables

Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$ , where $n$ denotes the number of stages and $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function. Then this type of optimization problems are called \emph{multi-stage problems}. Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being scalar and Let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$. The idea is that optimization can be done one stage at a time over $n$ stages after observing the state $X_{i}$ in each stage $i$. The value $\theta_{i}$ in stage $i$ has bearing on the cost of all subsequent stages $i+1,...,n$. This is  a problem of dynamic optimization. Approaches such as Dynamic programming can be used to solve this optimization problem.\par
Now we consider sub-class of multi-stage problems having infinite number of stages and having \emph{long-run average cost function} as objective. Objective function here is of the form:
\begin{align}\label{eq:long}
 f(\theta) = \lim_{n \to \infty} \frac{1}{n} ~ E \left[ \sum_{i=1}^{n} h_{i}(X_{i})\right ] ,
\end{align}
where $X_{i}$ is defined same as before and we assume it depends on parameter $\theta$. Objective function \eqref{eq:long} in most cases is not known analytically. In such cases it makes difficult for the search procedure to update the search parameter with out estimating the cost over infinite long trajectory. Another subclassifications of optimization problems is based on feasible region $\Theta$, see \eqref{eq:opt}.

@@

The rest of the thesis is organised as follows: In Chapter~\ref{sec:2rdsa-ih}, we describe the improved Hessian estimation scheme, which is incorporated into the  2RDSA algorithm from \cite{prashanth2015rdsa}. We also presented the theoretical results for the 2RDSA algorithm with  improved Hessian estimation and the results from numerical experiments. In Chapter~\ref{sec:2rdsa-gen}, we describe the generalised RDSA scheme for both the gradient and Hessian estimates and also present the theoretical convergence results as well. The work related to improving the Hessian estimation scheme 2SPSA-3 was presented in Chapter~\ref{sec:2spsa-3-ih} and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks were provided.

>>>>

Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$ , where $n$ denotes the number of stages and $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function. Then this type of optimization problems are called \emph{multi-stage problems}. Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being scalar and Let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$. The idea is that optimization can be done one stage at a time over $n$ stages after observing the state $X_{i}$ in each stage $i$. The value $\theta_{i}$ in stage $i$ has bearing on the cost of all subsequent stages $i+1,...,n$. This is  a problem of dynamic optimization. Approaches such as Dynamic programming can be used to solve this optimization problem.\par
Now we consider sub-class of multi-stage problems having infinite number of stages and having \emph{long-run average cost function} as objective. Objective function here is of the form:
\begin{align}\label{eq:long}
 f(\theta) = \lim_{n \to \infty} \frac{1}{n} ~ E \left[ \sum_{i=1}^{n} h_{i}(X_{i})\right ] ,
\end{align}
where $X_{i}$ is defined same as before and we assume it depends on parameter $\theta$. Objective function \eqref{eq:long} in most cases is not known analytically. In such cases it makes difficult for the search procedure to update the search parameter with out estimating the cost over infinite long trajectory. This kind of objective functions are very common in reinforcement learning applications. Another subclassifications of optimization problems is based on feasible region $\Theta$, see \eqref{eq:opt}.

@@

The rest of the thesis is organised as follows: In Chapter~\ref{sec:2rdsa-ih}, we describe the improved Hessian estimation scheme, which is incorporated into the  2RDSA algorithm from \cite{prashanth2015rdsa}. We also presented the theoretical results for the 2RDSA algorithm with  improved Hessian estimation and the results from numerical experiments. In Chapter~\ref{sec:2rdsa-gen}, we describe the generalised RDSA scheme for both the gradient and Hessian estimates and also present the theoretical convergence results as well. The work related to improving the Hessian estimation scheme 2SPSA-3 was presented in Chapter~\ref{sec:2spsa-3-ih}. In Chapter~\ref{sec:expts}, we provide simulation experiments for all the methods proposed in the previous chapters and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks and future directions were provided.
